\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-05-23 doc/paper.tex
mini change to structure

\bigskip
\begin{abstract}
Same-source analyses are a major part of a Forensic Toolmark Examiner's
job. Technological advances, such as profilometers and confocal
microscopy allow to measure 3D surfaces in a high-resolution digitized
form. Methods for matching marks for a variety of tools have been
studied in the literature. Analysis of these digitized markings require
the use of statistical methods which can quantify the scientific
mechanism of comparing markings and serve as basis for an error rate
calculation. Striation marks on bullets are made by impurities in the
barrel. The Chumbley score provides us with another approach in the
same-source assessment of bullet striation marks.
\end{abstract}

@@

\hypertarget{introduction-and-background}{%
\section{Introduction and
Background}\label{introduction-and-background}}

\hypertarget{motivation}{%
\subsection{Motivation}\label{motivation}}

@@

Analysis of these digitized markings require the use of statistical
methods which can quantify the scientific mechanism of comparing
markings and serve as basis for an error rate calculation.
\citet{manytoolmarks2} define a relative distance metric and use it as
similarity measure between two toolmarks. \citet{manytoolmarks1} extract
many small segments in the markings of two toolmarks and compare
similarity using a maximum Pearson correlation coefficient. The Chumbley
scoring method, first introduced by \citet{chumbley}, uses a similar but
more extensive framework based on a Mann-Whitney U test of the resulting
correlation coefficients. This approach is non-deterministic, because
segments are chosen randomly. \citep{hadler} make the score
deterministic for each pair of toolmarks by choosing segments for
comparison systematically. This approach also ensures independence
between segments of striae. In this paper, we are investigating the
applicability of the Chumbley scoring method by \citet{hadler} to assess
striation marks on bullet lands for same-source identification.

@@

Striation marks on bullets are made by impurities in the barrel. As the
bullet travels through the barrel, these imperfections leave
``scratches'' on the bullet surface (see top of \autoref{fig:rgl}).
Typically, only striation marks in the land engraved areas (LEAs) are
considered \citep{afte-article1992}. Bullet lands are depressed areas
between the grooves made by the rifling action of the barrel. Compared
to toolmarks made by screwdrivers striation marks on bullets are
typically much smaller, both in length and in width. Bullets also have a
curved cross-sectional topography.

@@

@@

Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of
0.272 for an optimized window size of 140 pixels -- which is
considerably higher than the error rates achieved on matching toolmarks.
Unfortunately, method CS1 also has a high rate of failed tests --
situations, in which the algorithm does not provide a result, due to the
way different-shift pairs are constructed. Algorithm CS2 is introduced
here as a remedy for failed tests by introducing an alternate version of
choosing different-shift pairs. The algorithm CS2 introduced here is
constructed in a way that achieves on average a ten-fold reduction in
the number of failures. While reducing the failure rate, the algorithm
also shows an increase in the power of the test. Type II error rates of
CS2 reach a minimum of 0.217 for an optimized window size of 130 pixels.
This increase in power of CS2 over CS1 should also apply to previous
studies on toolmarks. It would be interesting to see these results using
the adjusted algorithm. Unfortunately, none of the studies have made the
data publicly accessible.

>>>>

\bigskip
\begin{abstract}
Same-source analyses are a major part of a Forensic Toolmark Examiner's
job. Technological advances, such as profilometers and confocal
microscopy allow to measure 3D surfaces in a high-resolution digitized
form. Methods for matching marks for a variety of tools have been
studied in the literature. Analysis of these digitized markings require
the use of statistical methods which can quantify the scientific
mechanism of comparing markings and serve as basis for an error rate
calculation. In this paper, we are investigating the applicability of
the Chumbley scoring method by \citet{hadler} to assess striation marks
on bullet lands for same-source identification. The Chumbley score
provides us with another approach in the same-source assessment of
bullet striation marks.
\end{abstract}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

Analysis of these digitized markings require the use of statistical
methods which can quantify the scientific mechanism of comparing
markings and serve as basis for an error rate calculation.
\citet{manytoolmarks2} define a relative distance metric and use it as
similarity measure between two toolmarks. \citet{manytoolmarks1} extract
many small segments in the markings of two toolmarks and compare
similarity using a maximum Pearson correlation coefficient. The Chumbley
scoring method, first introduced by \citet{chumbley}, uses a similar but
more extensive framework based on a Mann-Whitney U test of the resulting
correlation coefficients. This approach is non-deterministic, because
segments are chosen randomly. \citep{hadler} make the score
deterministic for each pair of toolmarks by choosing segments for
comparison systematically. This approach also ensures independence
between segments of striae.

@@

In this paper, we are investigating the applicability of the Chumbley
scoring method by \citet{hadler} to assess striation marks on bullet
lands for same-source identification. Striation marks on bullets are
made by impurities in the barrel. As the bullet travels through the
barrel, these imperfections leave ``scratches'' on the bullet surface
(see top of \autoref{fig:rgl}). Typically, only striation marks in the
land engraved areas (LEAs) are considered \citep{afte-article1992}.
Bullet lands are depressed areas between the grooves made by the rifling
action of the barrel. Compared to toolmarks made by screwdrivers
striation marks on bullets are typically much smaller, both in length
and in width. Bullets also have a curved cross-sectional topography.

@@

{\textcolor{orange}{What are we investigating in this paper, and what is the structure for the paper?}}

\hypertarget{background}{%
\section{Background}\label{background}}

@@

Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of
0.272 for an optimized window size of 140 pixels -- which is
considerably higher than the error rates achieved on matching toolmarks,
but is similar to other single-feature methods proposed for bullet
matching. Unfortunately, method CS1 also has a high rate of failed tests
-- situations, in which the algorithm does not provide a result, due to
the way different-shift pairs are constructed. Algorithm CS2 is
introduced here as a remedy for failed tests by introducing an alternate
version of choosing different-shift pairs. The algorithm CS2 introduced
here is constructed in a way that achieves on average a ten-fold
reduction in the number of failures. While reducing the failure rate,
the algorithm also shows an increase in the power of the test. Type II
error rates of CS2 reach a minimum of 0.217 for an optimized window size
of 130 pixels. This increase in power of CS2 over CS1 should also apply
to previous studies on toolmarks. It would be interesting to see these
results using the adjusted algorithm. Unfortunately, none of the studies
have made the data publicly accessible.
